---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# molevolvR

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of molevolvR is to ...

## Installation

You can install the development version of molevolvR from [GitHub](https://github.com/) with:

``` r
# Install Bioconductor
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.19")

# Install Development Version of molevolvR
BiocManager::install("JRaviLab/molevol_scripts", ref = "begin-package-setup", subdir = "package/MolEvolvR", auth_token = '<PersonalAccessToken>')
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example, eval = F}
library(molevolvR)
## basic example code
```

## MolEvolvR: v2 & beyond!

### Companion Repos
- [MolEvol data](https://github.com/jravilab/molevol_data/)
- [MolEvolvR app](https://github.com/jravilab/molevolvr_app/)
- [Psp app](https://github.com/jravilab/psp_app/)
- [CompGenomeR](https://github.com/jravilab/compgenomer/)
- [MicroGenomeR](https://github.com/jravilab/microgenomer/)

### Where's what
- [Companion Shiny App](https://github.com/jravilab/molevolvr_app/) | [web-app](http://jravilab.org/molevovlr)
- [Upstream `sh/R` scripts](https://github.com/jravilab/molevol_scripts/tree/master/upstream_scripts)
- [Downstream `R` scripts](https://github.com/jravilab/molevol_scripts/tree/master/R)
- [Test data](https://github.com/jravilab/molevolvr_app/tree/master/TestData)
- [Other scripts](https://github.com/jravilab/molevol_scripts/tree/master/molevol_scripts)

### The Team
- [Jake Krol](//github.com/jakekrol) | 2022–Now
- [Joe Burke](//github.com/jburke11) | 2021–2022
- [Samuel Z Chen](//github.com/samuelzornchen) | 2019–2021
- [Lo Sosinski](//github.com/lsosinski) | 2019–2020
- [Janani Ravi](//github.com/jananiravi) | 2019–Now
