---
title: "MolEvolvr Help"
subtitle: "_MolEvolvr help docs and tutorials._"
output: html_document
#runtime: shiny
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fontawesome)
library(tidyverse); library(knitr)
#library(shiny); library(shinydashboard)
```

# MolEvolvR Help docs & Tutorials

This page will provide you with instructions on how to use the app to its fullest potential and helpful video tutorials. Videos will demonstrate how to navigate the application after retrieving your results.

## The UI: Workflow and Usage

### How can you use the app?

Characterization of proteins is crucial to understanding the molecular basis of fundamental cellular processes. Several computational tools have been developed to characterize protein sequence, structure, function, and phylogeny, and we have created a workflow to allow for the user to analyze proteins of interest at all of these levels using these tools. It begins by accepting accession numbers or FASTA files of proteins of interest, identifies homologs, determines function, and provides phyletic spreads. These analyses provide insight into the protein's purpose within organisms as well as detail how they have evolved. This is done in order to assist with providing an overview of their importance to a particular biological process, or survival itself.

*ABSTRACT*
A critical aspect of understanding the molecular basis of fundamental cellular processes, such as pathogenesis or energy acquisition, is a thorough characterization of the underlying proteins. Several individual computational tools have been developed to characterize proteins at the level of sequence, structure, and function, especially in the context of molecular evolution and phylogeny. However, there still lacks a comprehensive and consolidated workflow for applying all the relevant tools for a protein of interest. We present here the MolEvolvR webapp: a streamlined approach for analyzing proteins in the context of molecular evolution and phylogeny. Our workflow currently starts with one or more proteins of interest -- identifying homologs across the tree of life, followed by a detailed functional characterization in terms of their domain architectures, secondary structure elements, cellular localization -- in the context of phyletic spreads and phylogenetic analyses. This detailed workflow has been implemented in an easy-to-use web application that allows for automated and streamlined molecular evolution and phylogenetic analyses for any protein(s) of interest, including sequence-structure-function relationships as well as lineage-specific adaptations. 

*LS NOTES*
- END GOAL: Provide functional analysis & evolution of proteins, help people make new & exciting discoveries! Why? 
- Functional analysis of proteins will lead to the discovery of its use within organisms. It can lead to a better understanding of biological pathways and provide an idea of the structure of the protein. 

--> what is the end goal?

- similar proteins
- ancestry of proteins (where they com from, lineage specific, share features with related proteins in different kingdoms)
- exact same da present in other lineages, what could that mean?
  - lineage specific?
- biology of homologs, what more can they learn from the app than just blast or iprscan alone
- if exact same domains are popping up across lineages, what does that say about function?

- phyletic spreads and lineages
- what are the questions they can answer/analyses they can do?
- what is the value?

#### Use cases

[Phage-shock proteins (psp)](http://jravilab.shinyapps.io/psp-evolution)

in broad terms of what was done
- sample protein(s) -- Psp accessions, homologous and non-homologous
  - tested all analyses on these proteins
  - full analysis on non-homologous proteins
  - domain architecture 


MolEvolvR's functionality is comprised of 4 different types of analysis:

(i) Analysis via Domain Architecture, which allows for exploration of domain interaction and domain propensities;
(ii) Identification of thousands of Homologs; this allows us to see the similarities across various species
(iii) Phylogenetic analysis, which may draw upon information obtained through the other analysis to make inferences and visualizations pertaining to the phyletic spread of the target proteins and their homologs;
(iv) visualizing and analyzing results from other sources, *e.g.,* BLAST suite (BLASTP, BLASTCLUST), Interproscan, Multiple Sequence Alignment (MSA).

target audience, why they would use it and how to use the outputs, use cases


## Entering data into the app

### Accession Numbers and Fasta (full analysis)
To begin the full analysis, the accession numbers associated with your proteins of interest can be entered into the "Upload" tab. They should be given in a comma-separated list -- they may be copy/pasted or uploaded as a `.csv` file. You then click on the "Generate FASTA" button and FASTA sequences are retrieved from NCBI [link here]. Additionally, you may give FASTA sequences copy/pasted or as a `.fa` or `.fasta` file.

#### Multiple accessions/FASTA of homologs

`Skip the finding homologs step -- go directly to DA, phylogenetic analysis or only one or the other based on what they choose (checkbox)`

Alternatively, if the user already has a list of homologous proteins, they may enter the workflow post-homology search. Their protein data, in the form of accession numbers or FASTA sequences, is run through our molecular characterization and phylogenetic analysis processes. Additionally, the MolEvolvr webapp can use Clustal Omega, Clustal W, or MUSCLE to perform multiple sequence alignments on the proteins. If the user prefers to use a different alignment algorithm, they may enter their aligned FASTA Sequences that have been aligned using the algorithm of their choosing.

#### Unrelated accessions

`full suite of analysis on each -- each member appears in the dropdown`

A user can start at the beginning of our workflow needing only one or more accession numbers for their protein(s) of interest. Various combinations of the tools can then be selected and the accession numbers will be run through the corresponding procedures in the MolEvol workflow. For instance, the user could perform DeltaBLAST on their input, run the homology search, and then generate the domain architecture using InterProScan.

#### Related accessions aka homologs

### Finding homologs

`Queryable interface for homolog data`

Deltablast/psiblast + parameter options To begin the homolog search, a protein's FASTA sequence or accession number is given to the application. If an accession number is given, the app will search for the corresponding FASTA sequence. The FASTA file is then run through either DELTA-BLAST ref or PSI-BLAST ref, which can be chosen by the user. Both are versions of BLASTp ref, with a slight difference to each other: PSI-BLAST runs a BLASTp search and creates a Position-Specific Scoring Matrix (PSSM), which is then used to search the BLAST databases for more matches; DELTA-BLAST also uses PSSMs, but first searches preconstructed PSSMs and the CDD database first before continuing iterations. Once the BLAST homology search completes, it moves on to clustering with BLASTClust ref. Clustering homologs is based on sequence similarity (type of amino acid in the protein) and identity (exactly matching amino acid) to one another, putting proteins that are most similar to one another in a cluster. Cluster names are then added based on the most prominent domain in each group of proteins and the number of proteins in the cluster.

#### Customizing your analysis

------------------------------------------------------------------------

### BLAST outputs from the NCBI BLAST web-interface

Users may also enter the workflow with data obtained from a BLAST run. This data can be run through BLASTCLUST to complete the homology search. The phylogenetic analysis and domain architecture components can then be applied.

The MolEvolvR web-app allows you to input BLAST results that have been run externally on the BLAST web-server for your protein(s) of interest. To help us help you, we've put together a help doc to recommend useful parameters to modify prior to setting up your BLAST runs. Additionally, we'll help you with identifying the right format to download these results in to ensure compatibility with our web-app.

**Here are some instructions on uploading BLASTP results to our "Upload" tab:** <br>

#### STEP 1 Entering accession numbers/fasta sequences and choosing parameters

First, enter your accession number(s) or fasta sequence(s) into the "Enter Query Sequence" box. Choose either the non-redundant (nr) or reference proteins (refseq_proteins) and which algorithm to run. If you would like to further filter your results based on lineages (e.g., species, genus, family, kingdom), please enter the name/taxID accordingly and toggle the box for include/exclude those results in your search. 'blastp' is great for standard protein blast, if you know very little about your protein. If you are interested in identifying remote homologs, we suggest using the PSI-BLAST. DELTA-BLAST works very well if your protein has domains of interest. Creating a job title for the run is optional and for your personal use.

```{r out.width="60%", eval=F}
include_graphics()
```

Choose the number of maximum target sequences to look up. Our recommendation is 5,000--10,000 total target sequences, or hits to ensure maximum inclusion of homologs. Next, go down to the PSI/PHI/DELTABLAST box at the bottom and choose your threshold value, or e-value. We recommend using 1e-5 (also written as 1x10-5 or 0.00001). Double check your parameters in the box at the bottom of the page, then press the BLAST button.

```{r out.width="60%", eval=F}
include_graphics()
```

#### Step 2: Downloading BLAST results

Once the BLAST algorithm has finished running, there will be an option on the upper left of the screen to download results. Click on the "Download" button and select the "Hit Table(text)" option. Once these results are downloaded, you can directly upload these text files to the MolEvolvR web-app.

```{r out.width="60%", eval=F}
include_graphics()
```

```{r out.width="60%", eval=F}
include_graphics()
```

------------------------------------------------------------------------

### InterProScan outputs from the Iprscan5 web-interface

Running InterProScan

#### Step 1 Enter FASTA sequence

Input your protein's fasta sequence by copy/pasting into the box or upload the FASTA file via the "Choose file" button. Double check the sequence is valid with the green check mark next to it. Click the "Search" button to begin the search.

```{r out.width="60%", eval=F}
include_graphics()
```

#### Step 2 Downloading InterProScan results & uploading to MolEvolvR

Download the results in the TSV format under the "Export" dropdown menu. Upload the TSV file to the MolEvolvR app for visualization and further analysis of the domain architectures.

```{r out.width="60%", eval=F}
include_graphics()
```

------------------------------------------------------------------------

## Retrieving results from a prior run

After submitting proteins to the app, take a break! The full analysis takes around 4-5 hours to complete, domain architechture takes 1-2 hours, and just InterProScan takes about 30 minutes to complete. *UPDATE BASED ON BENCHMARKING TIMES, THIS IS JUST AN ESTIMATE BASED ON HOW LONG LAUREN HAS TO STARE AT HER COMPUTER*

You will receive and 8 character alphanumeric code before you close the application, which you will enter into the designated box under the Upload tab. Your results will then populate the screen. There are summary tables, figures, and interactive plots for you to play around with. If you require assistance navigating, please watch the tutorial at the end of this help page.

------------------------------------------------------------------------

/data/research/jravilab/molevolvr_app/scripts/ui/legendText.R <-- use as guide to explain

## Domain Architecture

A protein's domain architecture (DA) refers to the order of specific functional regions of a protein. Currently, InterProScan alone is used by MolEvolvR to characterize the domain architecture of a protein. 

`Iprscan; Add benchmarking times once that’s fully implemented, add tables/figures`

### Summary stats
A summary table is provided after completion of the analyses. This provides an overview of what domains are conserved across lineages. 

### Phyletic spreads
Several visualizations are provided to display the phylogeny of the proteins. Heatmaps are provided that display the occurrence of homologs within specific lineages. An interactive sunburst plot is created to show the phyletic spread of the query protein across bacteria, archaea, and eukaryota. Hovering over each section of the plot displays the lineage. 

### Upset plots
An upset plot shows the frequencies of domains and domain architectures within the homologs. This shows 

Blue histogram: Distribution of the constituent domains underlying all homologs.
Combination matrix: Various configurations of the constituent domains
that come together in the domain architectures.
Red histogram: Frequency of occurrences of the indicated domain architectures.

### DA visualizations
description

### Proximity network
description

------------------------------------------------------------------------

## Phylogeny

The dictionary defines phylogenisis as the evolutionary development and diversification of a species or group of organisms, or of a particular feature of an organism. Phylogenetic analysis of proteins provides key insights into their development and evolution. The conservation of certain portions through lineages or across domains of life could indicate the importance of the protein to certain biological processes.

### Phyletic spread of homologs

Query Heatmap, SunBurst -- based only on lineage
MolEvolvR creates an easy to read and 

### Multiple Sequence Alignments and Phylogenetic Trees
Tree -- based on amino acid sequence conservation

------------------------------------------------------------------------

## Tutorials

We have provided video tutorials of how to run and download your own web blast or InterProScan files, load data (accession numbers, fasta files, web blast, web InterProScan) into the app. These tutorial videos also demonstrate how to run your analyses within the app, how to access your data after it has been processed, how to navigate the app, and how to download your ready-to-publish figures!

------------------------------------------------------------------------

## Contact -- Twitter handles??
Janani Ravi       |   janani@msu.edu
Samuel Chen       |   chensam2@msu.edu
Lauren Sosinski   |   sosinsk7@msu.edu
