---
title: "JRaviLab"
subtitle: "Software and Data"
output: html_document
#runtime: shiny
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fontawesome)
#library(shiny); library(shinydashboard)
```

------------------------------------------------------------------------

# [`r fa("users", fill = "#384CB7")`](https://jravilab.github.io) [`r fa("github", fill = "#1FA67A")`](https://github.com/jravilab) [`r fa("twitter", fill = "#1DA1F2")`](https://twitter.com/jravilab) [`r fa("paper-plane", fill = "red")`](mailto:janani@msu.edu)

## *Computational Systems Biology of Infectious Diseases*

We develop discovery platforms that enable experimental and computational biologists use our approaches to study the molecular basis of pathogenesis and intervention of infectious diseases.

We welcome collaborations: if you are interested in studying an infectious disease/pathogen using our platforms or if you would like to see your computational tool incorporated into our platforms, do [get in touch](mailto:janani@msu.edu).

[**Learn more about our group here. `r fa("users", fill = "grey20")`**](https://jravilab.github.io)

------------------------------------------------------------------------

## Software & Methods
[MolEvolvR](#molevolvr) | [Connectivity Scores](#connectivity_scores)

### MolEvolvR: A web-app for molecular evolution and phylogeny {#molevolvr}

**MolEvolvR: A web-app for characterizing proteins using molecular evolution and phylogeny** <br>
*bioRxiv* 2022. doi: https://doi.org/10.1101/2022.02.18.461833. <br>
Joseph T Burke\*, Samuel Z Chen\*, Lo M Sosinski\*, John B Johnson, Janani Ravi^\#^.
<br> \*Co-primary authors. ^\#^Corresponding author: [janani\@msu.edu](mailto:janani@msu.edu) <br>
[`r fa("link", fill = "#1FA67A")`](http://jravilab.org/molevolvr) **Web-app:** [jravilab.org/molevolvr](http://jravilab.org/molevolvr)

```{r molevolvr, echo=FALSE, eval=F}
shinyAppDir(
  system.file("molevolvr", package = "shiny"),
  options = list(width = "100%", height = 550)
)
```

### Connectivity Scores for Drug Repurposing {#connectivity_scores}

**Reconciling Multiple Connectivity Scores for Drug Repurposing** <br>
*Brief Bioinform.* 2021 Nov 5;22(6):bbab161. doi: http://doi.org/10.1093/bib/bbab161. PMID: 34013329; PMCID: PMC8597919. <br> Kewalin Samart\*, Phoebe Tuyishime\*, Arjun Krishnan^\#^, Janani Ravi^\#^. <br>
\*Co-primary authors. ^\#^Corresponding authors: [arjun\@msu.edu](mailto:arjun@msu.edu); [janani\@msu.edu](mailto:janani@msu.edu) <br>
[`r fa("link", fill = "#1FA67A")`](https://jravilab.github.io/connectivity_score_review) **Live doc:** [jravilab.github.io/connectivity_score_review](https://jravilab.github.io/connectivity_score_review)

***

## Data repos & Web-apps

[Bacterial DciA](#dcia_evolution) | [Listeria Internalins](#inlp_evolution) | [Bacterial phage defense](#avcd_evolution) | [Phage shock stress response](#psp_evolution) 

### Bacterial DciA
DciA helicase operators exhibit diversity across bacterial phyla <br>
*bioRxiv* 2022.01.24.477630; doi: https://doi.org/10.1101/2022.01.24.477630 <br>
Helen C. Blaine\*, Joseph T. Burke\*, Janani Ravi^\#^, Christina L. Stallings^\#^ <br>
\*Co-primary authors. ^\#^Corresponding authors: [stallings\@wustl.edu](mailto:stallings@wustl.edu); [janani\@msu.edu](mailto:janani@msu.edu) <br>
[`r fa("link", fill = "#1FA67A")`](https://github.com/JRaviLab/dcia_evolution) **Data:** https://github.com/JRaviLab/dcia_evolution

### Internalins in Listeria {#inlp_evolution}
Novel Internalin P homologs in Listeria ivanovii londoniensis and Listeria seeligeri. <br>
*bioRxiv* 2022.01.19.476994; doi: https://doi.org/10.1101/2022.01.19.476994 (accepted at *Microbial Genomics*) <br>
Kayla N Conner\*, Joseph T Burke\*, Janani Ravi^\#^, Jonathan W Hardy^\#^ <br>
\*Co-primary authors. ^\#^Corresponding authors: [hardyjon\@msu.edu](mailto:hardyjon@msu.edu); [janani\@msu.edu](mailto:janani@msu.edu) <br>
[`r fa("link", fill = "#1FA67A")`](https://github.com/JRaviLab/inlp_listeria) **Data:** https://github.com/JRaviLab/inlp_listeria

### Bacterial phage defense {#avcd_evolution}
**A Broadly Conserved Deoxycytidine Deaminase Protects Bacteria from Phage Infection** <br>
*bioRxiv* 2021.03.31.437871; doi: https://doi.org/10.1101/2021.03.31.437871 (accepted at *Nature Microbiology*) <br>
Hsueh BY\*, Severin GB\*, Elg CA, Waldron EJ, Kant A, Wessel AJ, Dover JA, Rhoades CR, Ridenhour BJ, Parent KN, Neiditch MB, Ravi J, Top EM, and Waters CM. <br>
\*Co-primary authors. Corresponding author: watersc3@msu.edu <br>
[`r fa("link", fill = "#1FA67A")`](https://github.com/JRaviLab/phage_defense_avcd) **Data:** https://github.com/JRaviLab/phage_defense_avcd

### Phage shock protein stress response system {#psp_evolution}

**Phage-shock-protein (Psp) Envelope Stress Response: Evolutionary History & Discovery of Novel Players** <br>
*bioRxiv* 2020. doi: <https://doi.org/10.1101/2020.09.24.301986> <br>
Janani Ravi^\#^, Vivek Anantharaman, Samuel Zorn Chen, Pratik Datta, L Aravind^\#^, Maria Laura Gennaro^\#^. <br>
^\#^Corresponding authors: [janani\@msu.edu](mailto:janani@msu.edu); [aravind\@nih.gov](mailto:aravind@nih.gov); [marila.gennaro\@rutgers.edu](mailto:marila.gennaro@rutgers.edu). <br>
[`r fa("link", fill = "#1FA67A")`](https://jravilab.shinyapps.io/psp-evolution) **Web-app:** [jravilab.shinyapps.io/psp-evolution](https://jravilab.shinyapps.io/psp-evolution). Web-app by JR and SZC.

```{r psp-evolution, echo=FALSE, eval=F}
shinyAppDir(
  system.file("psp-evolution", package = "shiny"),
  options = list(width = "100%", height = 550)
)
```
***
## Other
**Ten simple rules to host an inclusive conference.** <br>
2022 (accepted at *PLoS Computational Biology*) <br>
Rocío Joo\*^\#^, Andrea Sánchez-Tapia\*, Sara Mortara, Yanina B Saibene, Heather Turner, Dorothea H Peter, Natalia S Morandeira, Matthias Bannert, Batool Almazrouq, Elizabeth Hare, Laura Ación, Juan P Narváez-Gómez, Marcela A Córdoba, Federico Marini, Rita Giordano, Silvia Canelón, Anicet Ebou, Adithi Upadhya, Joselyn Chávez, Janani Ravi^\#^. <br>
\*Co-primary authors. ^\#^Corresponding authors: [rocio.joo\@globalfishingwatch.org](mailto:rocio.joo@globalfishingwatch.org); [janani\@msu.edu](mailto:janani@msu.edu) <br>
[`r fa("link", fill = "#1FA67A")`](https://github.com/rociojoo/Ten-simple-rules-inclusive-conference) **Repo:** [github.com/rociojoo/Ten-simple-rules-inclusive-conference](https://github.com/rociojoo/Ten-simple-rules-inclusive-conference).
