---
title: "JRaviLab"
subtitle: "Web-applications and Software"
output: html_document
runtime: shiny_prerendered
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fontawesome)
library(shiny)
library(shinydashboard)
```

------------------------------------------------------------------------

# [`r fa("users", fill = "#384CB7")`](https://jravilab.github.io) [`r fa("github", fill = "#1FA67A")`](https://github.com/jravilab) [`r fa("twitter", fill = "#1DA1F2")`](https://twitter.com/jravilab) [`r fa("paper-plane", fill = "red")`](mailto:janani@msu.edu)

## *Computational Systems Biology of Infectious Diseases*

We are always working on developing discovery platforms that enable experimentalists and other computational scientists use our approaches to study the molecular basis of pathogenesis and intervention of infectious diseases.

We also welcome collaborations. If you are interested in studying an infectious disease/pathogen using our platforms or if you would like to see your computational tool incorporated into our platforms, do [get in touch](mailto:janani@msu.edu).

[**Learn more about our group here. `r fa("users", fill = "grey20")`**](https://jravilab.github.io)

------------------------------------------------------------------------

## Web-apps \| Software \| Code \| Methods

-   [MolEvolvR](#molevolvr)
-   [Psp Evolution](#psp_evolution)
-   [Connectivity Scores](#connectivity_scores)

### MolEvolvR: A web-app for molecular evolution and phylogeny {#molevolvr}

#### [`r fa("link", fill = "#1FA67A")`](jravilab.org/molevolvr) [jravilab.org/molevolvr](jravilab.org/molevolvr)

Samuel Zorn Chen, Lauren Marie Sosinski, John Johnston, Janani Ravi. <br> Corresponding author: [janani\@msu.edu](mailto:janani@msu.edu){.email}

```{r molevolvr, echo=FALSE, eval=T}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"), #molevolvr
  options = list(
    width = "100%", height = 550
  )
)
```

### Phage shock protein stress response system {#psp_evolution}

#### [`r fa("link", fill = "#1FA67A")`](jravilab.shinyapps.io/psp-evolution) [jravilab.shinyapps.io/psp-evolution](jravilab.shinyapps.io/psp-evolution)

**Phage-shock-protein (Psp) Envelope Stress Response: Evolutionary History & Discovery of Novel Players** <br> Janani Ravi\*, Vivek Anantharaman, Samuel Zorn Chen, Pratik Datta, L Aravind\*, Maria Laura Gennaro\*. \*Corresponding authors: [janani\@msu.edu](mailto:janani@msu.edu){.email}; [aravind\@nih.gov](mailto:aravind@nih.gov){.email}; [marila.gennaro\@rutgers.edu](mailto:marila.gennaro@rutgers.edu){.email}. Web-app by JR and SZC. <br> [**bioRxiv 2020**](https://doi.org/10.1101/2020.09.24.301986)

```{r psp-evolution, echo=FALSE, eval=F}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"), #psp-evolution
  options = list(
    width = "100%", height = 550
  )
)
```

### Connectivity Scores for Drug Repurposing {#connectivity_scores}

#### [`r fa("link", fill = "#1FA67A")`](jravilab.github.io/connectivity_score_review) [jravilab.shinyapps.io/psp-evolution](jravilab.github.io/connectivity_score_review)

**Reconciling Multiple Connectivity Scores for Drug Repurposing** <br> Kewalin Samart\*, Phoebe Tuyishime\*, Arjun Krishnan^\#^, Janani Ravi^\#^. <br> \*These authors contributed equally to this work. ^\#^Corresponding authors: [arjun\@msu.edu](mailto:arjun@msu.edu){.email}; [janani\@msu.edu](mailto:janani@msu.edu){.email} <br> [**arXiv 2020**](https://arxiv.org/abs/2009.09317)
